---
title: "In-One-Chart"
author: "Luiza Swierzawska"
date: "2026-02-12"
output: html_document
---

I collected the data from https://www.nursetogether.com/nurse-strikes/. I double-checked the information with news articles online and compiled a list including the duration of each strike. The list covers strikes from years 2020-2025.
```{r database, echo=TRUE}
nurse_strikes <- read.csv("nurse-strikes-db.csv")
```

The current database is missing strikes from this year which are the main focus of the story. Below I added the rows with NYC strikes. 
```{r database, echo=TRUE}
 
# Remove the extra X columns
nurse_strikes <- nurse_strikes[, c("name", "state", "dates", "duration_days")]

# Now add the new rows
new_rows <- data.frame(
  name = c("NewYork-Presbyterian", 
           "Montefiore Medical Center, NYC", 
           "Mount Sinai Hospital, NYC"),
  state = c("New York", "New York", "New York"),
  dates = c("January 12 - ongoing 2026", 
            "January 12 - February 9 2026", 
            "January 12 - February 9 2026"),
  duration_days = c(31, 29, 29),
  stringsAsFactors = FALSE
)

nurse_strikes <- rbind(nurse_strikes, new_rows)
```

```{r}
# Sort by duration_days in descending order
nurse_strikes_sorted <- nurse_strikes[order(-nurse_strikes$duration_days), ]

# Display top 20
head(nurse_strikes_sorted, 20)
```
In order to highlight the longest nurses' strikes in the recent years in the US and to not make the visualization too overwhelming, I've decided to include only the strikes that lasted longer than 2 weeks. 

```{r}
library(ggplot2)
library(dplyr)

long_strikes <- nurse_strikes %>%
  filter(duration_days >= 14) %>%
  mutate(
    is_nyc = grepl("NYC|NewYork-Presbyterian|Mount Sinai Hospital, NYC|Montefiore Medical Center, NYC", name),
    is_ongoing = grepl("ongoing", dates, ignore.case = TRUE)
  ) %>%
  arrange(duration_days)

ggplot(long_strikes, aes(y = reorder(name, duration_days))) +
  
  # Lines for all strikes
  geom_linerange(
    aes(xmin = 0, xmax = duration_days, color = is_nyc),
    linewidth = 1.4
  ) +
  
  # Points for completed strikes only
  geom_point(
    data = filter(long_strikes, !is_ongoing),
    aes(x = duration_days, color = is_nyc),
    size = 3
  ) +
  
    annotate("text", 
           x = 35, 
           y = "NewYork-Presbyterian", 
           label = "Ongoing", 
           hjust = 0, 
           vjust = 0.5,
           size = 3.5, 
           color = "red",
           fontface = "italic") +
  
  geom_segment(
    data = filter(long_strikes, is_ongoing),
    aes(x = 0, xend = duration_days, y = name, yend = name),
    arrow = arrow(length = unit(0.3, "cm"), type = "closed"),
    color = "red",
    linewidth = 1.4
  ) +
  scale_color_manual(
    values = c("TRUE" = "red", "FALSE" = "grey60"),
    labels = c("TRUE" = "2026 NYC Strikes", "FALSE" = "Other US Strikes"),
    name = NULL
  ) +
  
  scale_x_continuous(breaks = seq(0, 350, 30)) +  # Grid every 30 days
  
  labs(
    title = "2026 New York nurses' strike is among the longest\nin recent US history",
    subtitle = "When compared to Nurses Together's compilation of strikes from 2020-2025,the 2026 NYC strike ranks among the longest in recent history.\nNew York State Nurses Association went on strike on January 12, with NewYork-Presbyterian still ongoing.",
    x = "Duration (days)",
    y = NULL
  ) +
  
  theme_minimal(base_family = "sans") +
  
  theme(
    plot.title.position = "plot",
    plot.title = element_text(size = 18, face = "bold", hjust = 0),
    plot.subtitle = element_text(size = 9, hjust = 0, margin = margin(b = 15)),
    axis.text.y = element_text(size = 9, hjust = 0),
    axis.title.x = element_text(size = 7),
    legend.position = "top",
    legend.justification = "left",
    legend.text = element_text(size = 10),
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(15, 15, 15, 15)
  )
```
TK - still need to figure out the dates so it's not to overwhelming. Probably will do something like: Hospital Name (Year), State
TK - I have interviews with the nurses who were protesting in front of the NYC Presbyterian. I just need to figure out how to nicely put it together. Still looking for experts.



